---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arpctheme

<!-- badges: start -->
<!-- badges: end -->

The `arpctheme` package provides a ggplot2 theme designed specifically for ARPC data visualizations. The package also contains an export function that allows users to consistently export figures in a standardized formatting. 

## Installation

You can install the development version of arpctheme from GitHub with:

```r
# install.packages("pak")
pak::pak("dylan-turner25/arpctheme")
```

## Quick Start


Below is a standard ggplot scatter plot with out the arpc theme applied. 
```{r example}
library(ggplot2)

# Basic usage a standard ggplot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "A standard ggplot",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders")
```

Adding the `+ theme_arpc()` function applies the ARPC theme.
```{r example-theme}
# Load the arpctheme package
library(arpctheme)
# Apply the ARPC theme
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc()
```
### Adding a Logo

Further, an ARPC logo can be added using `+ logo()`. The logo will automatically be placed in the bottom right corner of the plot, but can be customized to appear in other locations. The logo position can be changes by specifying the `position` argument, which can be set to one of the following: "bottom-right", "bottom-left", "top-right", "top-left", "bottom", "top", "left", or "right". These text based position arguments will often end up leaving you disappointed with the placement of the logo. In such case, the logo can also be placed in a specific location using coordinates, where (0,0) is the bottom left corner of the plot and (1,1) is the top right corner. Setting coordinates outside the 0 to 1 range is also possible and will place the logo into the margins of the plot. The size of the logo can be adjusted using the `size` argument, which defaults to 0.3. The `alpha` argument can be used to adjust the transparency of the logo to create a watermark type effect. 
```{r example-logo}

# Add the ARPC logo to the plot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot with Logo",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc() +
  logo()
```
Example: place the logo in the upper right hand corner and increase the size.
```{r example-logo2}
# Add the ARPC logo to the plot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot with Logo",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc() +
  logo(position = c(.75,1.05), size = 0.5)
```

Example: Use the logo as a watermark in the background of the plot.
```{r timeseries-example}
# some made up data
crop_data <- data.frame(
  year = rep(2020:2024, 4),
  yield = c(85, 88, 92, 89, 94,   # Corn
           45, 48, 52, 50, 55,    # Soybeans  
           65, 68, 71, 69, 75, # Wheat
           50,50,50,75,90),   # Rice
      
  crop = rep(c("Corn", "Soybeans", "Wheat","Rice"), each = 5)
)

ggplot(crop_data, aes(x = year, y = yield, color = crop)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  theme_arpc() +
  logo(position = c(.5,.7), size = 1, alpha = .2) +
  labs(
    title = "Made Up Crop Yields Over Time",
    subtitle = "",
    x = "Year", 
    y = "Yield (bushels/acre)",
    color = "Crop Type"
  )
```

## NDSU Color Pallete

The package automatically applies colors from the [NDSU brand guidlines](https://www.ndsu.edu/sites/default/files/fileadmin/www.ur.ndsu.edu/NDSU-Brand-Guidelines-040125.pdf):

1. **NDSU Green** (#00583d) - Primary brand color
2. **NDSU Yellow** (#FFC425) - Secondary brand color  
3. **Rust** (#B83E27) - Third priority
4. **Night** (#0F374B) - Fourth priority
5. Additional colors: Teal, Sage, Dark Green, etc.

### Access Individual Colors

```{r colors-example}
# return the hex codes for all ndsu colors
ndsu_colors()

# get the hex code for just ndsu green
ndsu_colors("green")

# get the hex codes for ndsu green and ndsu yellow
ndsu_colors(c("green", "yellow"))
```

