---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arpctheme

<!-- badges: start -->
<!-- badges: end -->

The `arpctheme` package provides a comprehensive ggplot2 theme and color system designed specifically for ARPC (Agricultural Research and Policy Center) data visualizations. It ensures consistent, professional styling that incorporates NDSU brand colors and typography.

## Installation

You can install the development version of arpctheme from GitHub with:

```r
# install.packages("pak")
pak::pak("dylan-turner25/arpctheme")
```

## Quick Start


Below is a standard ggplot scatter plot with out the arpc theme applied. 
```{r example}
library(ggplot2)

# Basic usage a standard ggplot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "A standard ggplot",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders")
```

Adding the `+ theme_arpc()` function applies the ARPC theme.
```{r example-theme}
# Load the arpctheme package
library(arpctheme)
# Apply the ARPC theme
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc()
```
### Adding a Logo

Further, an ARPC logo can be added using `+ logo()`. The logo will automatically be placed in the bottom right corner of the plot, but can be customized to appear in other locations. The logo position can be changes by specifying the `position` argument, which can be set to one of the following: "bottom-right", "bottom-left", "top-right", "top-left", "bottom", "top", "left", or "right". These text based position arguments will often end up leaving you disappointed with the placement of the logo. In such case, the logo can also be placed in a specific location using coordinates, where (0,0) is the bottom left corner of the plot and (1,1) is the top right corner. Setting coordinates outside the 0 to 1 range is also possible and will place the logo into the margins of the plot. The size of the logo can be adjusted using the `size` argument, which defaults to 0.3. 
```{r example-logo}

# Add the ARPC logo to the plot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot with Logo",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc() +
  logo()
```
Example: place the logo in the upper right hand corner and increase the size.
```{r example-logo2}
# Add the ARPC logo to the plot
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point(size = 2) +
  labs(title = "ARPC Themed ggplot with Logo",
       x = "Weight (1000 lbs)", 
       y = "Miles per Gallon",
       color = "Cylinders") +
  theme_arpc() +
  logo(position = c(.75,1.05), size = 0.5)
```




## NDSU Color System

The package automatically applies NDSU brand colors in priority order:

1. **NDSU Green** (#00583d) - Primary brand color
2. **NDSU Yellow** (#FFC425) - Secondary brand color  
3. **Rust** (#B83E27) - Third priority
4. **Night** (#0F374B) - Fourth priority
5. Additional colors: Teal, Sage, Dark Green, etc.

### Access Individual Colors

```{r colors-example}
# Get all NDSU colors
ndsu_colors()

# Get specific colors
ndsu_colors("green")
ndsu_colors(c("green", "yellow", "rust"))
```



## Examples

### Time Series Plot

```{r timeseries-example}
# some made up data
crop_data <- data.frame(
  year = rep(2020:2024, 3),
  yield = c(85, 88, 92, 89, 94,   # Corn
           45, 48, 52, 50, 55,    # Soybeans  
           65, 68, 71, 69, 75),   # Wheat
      
  crop = rep(c("Corn", "Soybeans", "Wheat"), each = 5)
)

ggplot(crop_data, aes(x = year, y = yield, color = crop)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  theme_arpc() +
  logo(position = c(.5,.5), size = 1, alpha = .3) +
  labs(
    title = "Made Up Crop Yields Over Time",
    subtitle = "",
    x = "Year", 
    y = "Yield (bushels/acre)",
    color = "Crop Type"
  )
```

### Bar Chart

```{r barchart-example}
ggplot(mtcars, aes(x = factor(cyl), fill = factor(cyl))) +
  geom_bar() +
  theme_arpc() +
  logo(position = "bottom-right", size = 0.5) +
  labs(
    title = "Distribution by Cylinder Count",
    x = "Cylinders",
    y = "Count",
    fill = "Cylinders"
  )
```

## Customization

### Override Colors

Users can override the automatic NDSU colors when needed:

```{r override-example, eval=FALSE}
ggplot(data, aes(x, y, color = group)) +
  geom_point() +
  theme_arpc() +
  scale_color_manual(values = c("red", "blue", "green"))  # Overrides NDSU colors
```

### Font Customization

```{r font-example, eval=FALSE}
# Use different font family
ggplot(data, aes(x, y)) +
  geom_point() +
  theme_arpc(base_family = "Arial")

# Adjust font size
ggplot(data, aes(x, y)) +
  geom_point() +
  theme_arpc(base_size = 16)
```

## Best Practices

1. **Consistent Usage**: Always use `theme_arpc()` for ARPC publications to ensure brand consistency
2. **Logo Placement**: Choose logo positions that don't interfere with data visualization (automatic clipping ensures logos appear correctly)
3. **Color Priority**: Let the automatic color system handle color assignment for consistency
4. **Export Standards**: Use `export_arpc()` with consistent dimensions for publication-ready figures
5. **Font Considerations**: The default CMR font works well for academic publications

## Contributing

This package is designed specifically for ARPC research visualization needs. For issues or suggestions, please contact the package maintainer.

## License

MIT License - see LICENSE file for details.